---
title: "JellyfishPlotting"
author: "Caroline Newell"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries and data

```{r}
library(tidyverse)
library(readxl)
SMFS_09192023 <- read_csv("C:/Users/cnewe/OneDrive/Documents/WFC 98/Guest Lecture/WFC98/SMFS_09192023.csv") #loading the data table we created.
```

# Calculating CPUE
We need to standardize our counts.

We only have positive instances of catch but we also want to include instances of zero catch.
We will do this by pivoting wider and filling with zeros.

Before we do so, we need to double check our new dataframe for any oddities (mostly concerned about NA's) that could have artificially arisen from our previous coding.



```{r}
glimpse(SMFS_09192023)

 
summary(SMFS_09192023$TowDuration) #this will be what we use to standardize our catch. The effort will be the duration of our sampling effort. We could also choose things like standardizing by number of sampling events, but in this case there are different lengths of sampling events so not all events have equal effort. So we standardize by trawl minute.

#We have NA's! But why??
```

## Investigating NAs

```{r}
investigating<-SMFS_09192023 %>% filter(is.na(TowDuration)) #Let's keep all rows with our NA tow duration.
n_distinct(investigating$SampleRowID) #How many sampling events have NA tow duration?
unique(investigating$TrawlComments) #Maybe the comments can tell us why we have these NAs?

unique(investigating$OrganismCode) #Is our species of interest contained within the NA dataset?
#while checking for Maeotias (the code for black sea jellyfish) we find an odd code "NO TRAWL"
InvestigatingFurther<-investigating %>% filter(OrganismCode=="NOTRAWL") #Let's see if all of our sampling events with NA tow duration were actually not trawls?
n_distinct(InvestigatingFurther$SampleRowID)

```
Okay, we now know there are 19 trawls which are represented in our data that actually weren't trawls. But we know 20 samples had NA tow duration, so there is one trawl that is missing tow duration information.
To be absolutely sure this is from the original data, we can look back at that sample row ID in the raw Trawl Effort data we downloaded. We can also go and find the original physical datasheet to verify duration wasn't written down.
Since this appears to be an issue with the raw data, this is simply missing information that we want to filter out before calculating CPUE. Missing information is commonplace. We want to remove this before calculating CPUE since otherwise we would end up with NAN trying to divide counts by an NA value.

## Remove NAs
```{r}
SMFS_09192023<-SMFS_09192023 %>% filter(!is.na(TowDuration)) #when we include an ! before the filtered item, that means we want to remove it.
#is.na() is a useful argument which identifies NA values in the dataset.
```

 ## Pivot!
```{r}

unique(SMFS_09192023$OrganismCode) #these all will be turned into columns and we will fill their cells with counts for each trawl.
Wide_Table<-SMFS_09192023 %>% pivot_wider(names_from = OrganismCode, values_from = Count, values_fill=0)
glimpse(Wide_Table)

#We ended up with over 100 columns! Let's simplify for only columns of interest
JellyTable<-Wide_Table %>% select(SampleRowID, TowDuration, WaterTemperature, Salinity, DO, Secchi, MAEOTIAS)
#Add a column for CPUE
JellyTable$CPUE<-JellyTable$MAEOTIAS/JellyTable$TowDuration
summary(JellyTable$CPUE)#Looks reasonable. We see that most of the time we aren't catching jellyfish but when we do, its in huge numbers.
summary(JellyTable$WaterTemperature) #looking at our water quality parameters as well now
summary(JellyTable$Salinity)
summary(JellyTable$DO)
summary(JellyTable$Secchi)

```


# Plotting!
Let's make some exploratory plots now!
We don't have to worry about making them too fancy. At this stage it is just for us to look at, not for publication or presentation. 

```{r}
Temp<-JellyTable %>% 
  ggplot(aes(x=WaterTemperature, y=CPUE)) +
  geom_col() + 
  theme_classic() +
  labs(title = "Count per minute trawl by water temperature (C)")
Temp
  
Salinity<-JellyTable %>% 
  ggplot(aes(x=Salinity, y=CPUE)) +
  geom_col() + 
  theme_classic() +
  labs(title = "Count per minute trawl by salinity (ppt)")
Salinity

DO<-JellyTable %>% 
  ggplot(aes(x=DO, y=CPUE)) +
  geom_col() + 
  theme_classic() +
  labs(title = "Count per minute trawl by dissolved oxygen (mg/L)")
DO

Secchi<-JellyTable %>% 
  ggplot(aes(x=Secchi, y=CPUE)) +
  geom_col() + 
  theme_classic() +
  labs(title = "Count per minute trawl by secchi depth (cm)")
Secchi
library(cowplot) #needed for the next line
plot_grid(Salinity, Temp, DO, Secchi, nrow=2) 
```
